---
title: "Applied Statistical Programming - Spring 2022"
output: pdf_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{center}
{\Large{\textbf{Problem Set 3}}} \\
\vspace{4 bp}
Due Wednesday, March 16, 10:00 AM (Before Class) \\
\end{center}

\section*{Instructions}
\begin{enumerate}
  \item The following questions should each be answered within an Rmarkdown file. Be sure to provide many comments in your code blocks to facilitate grading. Undocumented code will not be graded.
  \item Work on git. Continue to work in the repository you forked from \url{https://github.com/johnsontr/AppliedStatisticalProgramming2022} and add your code for Problem Set 4. Commit and push frequently. Use meaningful commit messages because these will affect your grade.
  \item You may work in teams, but each student should develop their own Rmarkdown file. To be clear, there should be no copy and paste. Each keystroke in the assignment should be your own.
  \item For students new to programming, this may take a while. Get started.
\end{enumerate}

\section*{\texttt{tidyverse}}

Your task in this problem set is to combine two datasets in order to observe how many endorsements each candidate received using only \texttt{dplyr} functions. Use the same Presidential primary polls that were used for the in class worksheets on February 28 and March 2.

```{r}
# Change eval=FALSE in the code block. Install packages as appropriate.
#install.packages("fivethirtyeight")
library(fivethirtyeight)
library(tidyverse)
# URL to the data that you've used.
url <- 'https://jmontgomery.github.io/PDS/Datasets/president_primary_polls_feb2020.csv'
polls <- read_csv(url)
Endorsements <- endorsements_2020 # from the fiverthirtyeight package
```

First, create two new objects \texttt{polls} and \texttt{Endorsements}. Then complete the following.
\begin{itemize}
  \item Change the \texttt{Endorsements} variable name endorsee to \texttt{candidate\_name}.
  \item Change the \texttt{Endorsements} dataframe into a \texttt{tibble} object.
  \item Filter the \texttt{poll} variable to only include the following 6 candidates: Amy Klobuchar, Bernard Sanders,Elizabeth Warren, Joseph R. Biden Jr., Michael Bloomberg, Pete Buttigieg \textbf{and} subset the dataset to the following five variables: \texttt{candidate\_name, sample\_size, start\_date, party, pct}
  \item Compare the candidate names in the two datasets and find instances where the a candidates name is spelled differently i.e. Bernard vs. Bernie. Using only \texttt{dplyr} functions, make these the same across datasets. 
  \item Now combine the two datasets by candidate name using \texttt{dplyr} (there will only be five candidates after joining).
  \item Create a variable which indicates the number of endorsements for each of the five candidates using \texttt{dplyr}.
  \item Plot the number of endorsement each of the 5 candidates have using \texttt{ggplot()}. Save your plot as an object \texttt{p}.
  \item Rerun the previous line as follows: \texttt{p + theme\_dark()}. Notice how you can still customize your plot without rerunning the plot with new options.
  \item Now, using the knowledge from the last step change the label of the X and Y axes to be more informative, add a title. Save the plot in your forked repository.
\end{itemize}

```{r}
#Change variable name, using rename function and store it in our original df
Endorsements <- Endorsements %>% rename(candidate_name = endorsee)

#Change to a tibble and check the class
Endorsements <- Endorsements %>% as_tibble()
class(Endorsements)

#Filter to the candidates with an in command of a vector of the candidates of interest
polls <- polls %>% 
  filter(polls$candidate_name %in% c("Amy Klobuchar","Bernard Sanders", "Elizabeth Warren", "Joseph R. Biden Jr.", "Michael Bloomberg", "Pete Buttigieg")) %>% 
  select(candidate_name, sample_size, start_date, party, pct)
#build a table based on the candidate name to confirm only those 6 remain in our new object
table(polls$candidate_name)
#use summary to see that there are only 5 remaining variables as told to do
summary(polls)

#Check spelling of names in our Endorsements dataset
table(Endorsements$candidate_name)
#Here we can tell Amy Klobuchar, Elizabeth Warren, and Pete Buttigieg are spelled the same. Biden is "Joe Biden" here, while "Joseph R. Biden Jr." in our polls set, Sanders is "Bernie Sanders" here instead of "Bernard Sanders", and Bloomberg does not seem to exist in the endorsements set at all. 

#Now we need to change these using recode
Endorsements <- Endorsements %>% 
  mutate(candidate_name=recode(candidate_name, "Bernie Sanders" = "Bernard Sanders", "Joe Biden" = "Joseph R. Biden Jr."))

#Combine them into a dataset using an inner join by candidate name, then verify only the 5 expected remain
pe_combined <- polls %>% 
  inner_join(Endorsements, by="candidate_name")
table(pe_combined$candidate_name)

#Count the number of endorsements for each of the five
wide_pe_combined <- pe_combined %>% 
  group_by(candidate_name) %>% 
  summarize(length(points))

p <- ggplot(data = wide_pe_combined) + geom_col(aes(candidate_name,`length(points)`))

p + theme_dark()

p + labs(y = "Number of Endorsements", x = "Candidate", title = "2020 Democratic Presidential Candidate Endorsements")
ggsave("candidate_endorsements_2020.pdf",
       plot = last_plot())

```


\section*{Text-as-Data with \texttt{tidyverse}}

For this question you will be analyzing Tweets from President Trump for various characteristics. Load in the following packages and data:

```{r}
# Change eval=FALSE in the code block. Install packages as appropriate.
library(tidyverse)
#install.packages('tm')
library(tm) 
#install.packages('lubridate')
library(lubridate)
#install.packages('wordcloud')
library(wordcloud)
library(knitr)
library(broom)
library(tidytext)
trump_tweets_url <- 'https://politicaldatascience.com/PDS/Datasets/trump_tweets.csv'
tweets <- read_csv(trump_tweets_url)
```

\begin{itemize}
  \item First separate the \texttt{created\_at} variable into two new variables where the date and the time are in separate columns. After you do that, then report the range of dates that is in this dataset.
  \item Using \texttt{dplyr} subset the data to only include original tweets (remove retweets) and show the text of the President's \textbf{top 5} most popular and most retweeted tweets. (Hint: The \texttt{match} function can help you find the index once you identify the largest values.) 
  \item Create a \textit{corpus} of the tweet content and put this into the object \texttt{Corpus} using the \texttt{tm} (text mining) package. (Hint: Do the assigned readings.)
  \item Remove extraneous whitespace, remove numbers and punctuation, convert everything to lower case and remove 'stop words' that have little substantive meaning (the, a, it).
  \item Now create a \texttt{wordcloud} to visualize the top 50 words the President uses in his tweets. Use only words that occur at least three times. Display the plot with words in random order and use 50 random colors. Save the plot into your forked repository.\item Create a \textit{document term matrix} called \texttt{DTM} that includes the argument \texttt{ control = list(weighting = weightTfIdf)}
  \item Finally, report the 50 words with the the highest tf.idf scores using a lower frequency bound of .8.
\end{itemize}

```{r}
#library a package necessary for date conversion
library(anytime)
#split the data using mutate and str_split
created_split <- tweets %>% mutate(date_created = str_split(created_at, pattern = " "))
#convert it to the correct col/row orientation
time <- t(as.data.frame(created_split$date_created))
#name columns
colnames(time) <- c("date","time")
#combine into our original dataset
tweets <- cbind(tweets,time)
#convert to date format for finding the oldest and newest
tweets$date <- anydate(tweets$date)
#oldest tweet day
min(tweets$date)
#newest tweet day
max(tweets$date)
#So the range of dates is 1/1/2014 to 2/14/2020

#Now to subset only to tweets
tweets_only <- tweets %>% filter(is_retweet == FALSE)
#verify our result
table(tweets_only$is_retweet)
#get the most popular
most_liked <- tweets_only %>% 
  slice_max(favorite_count, n=5)
most_liked$text
#Get the most retweeted
most_retweeted <- tweets_only %>% 
  slice_max(retweet_count, n=5)
most_retweeted$text

#Now for the corpus
#Create the corpus
Corpus <- VCorpus(VectorSource(tweets_only$text))
#check that it is working
writeLines(head(strwrap(Corpus[[420]]),10))

#Now to clean the tweets; I'm splitting it for processing power, but keeping some together for efficiency
#first I'll remove all URLs
removepattern <- content_transformer(function(x,pattern){
  return(gsub(pattern,"",x))
})

Corpus <- tm_map(Corpus, removepattern, "?(f|ht)(tp)(s?)(://)(.*)(.|/])(.*)")

#Next, I will make transform everything to lower case, remove whitespace, and stopwords
Corpus <- Corpus %>% 
  tm_map(content_transformer(tolower)) %>% 
  tm_map(stripWhitespace) %>% 
  tm_map(removeWords, stopwords("english")) 

#next I'll remove punctuation and numbers
Corpus <- Corpus %>% 
  tm_map(removePunctuation) %>% 
  tm_map(removeNumbers)

#Test 
writeLines(head(strwrap(Corpus[[420]]),10))
#Make the word cloud with random color and order
set.seed(125)
wc <- wordcloud(Corpus, min.freq = 3, 
                random.order = T,
                random.color = T,
                max.words = 50,
                colors = sample(colors()))

#Document Term matrix
dtm <- DocumentTermMatrix(Corpus, control = list(weighting = weightTfIdf, global=c(0.8,Inf)))

#Get the top 50:
#Start by turning it into a data frame
dtm <- tidy(dtm)
#calculate the scores
dtm_tfidf <- bind_tf_idf(dtm, term = term, document = document, n = count)
top50_terms <- dtm_tfidf %>% 
  group_by(term) %>% 
  filter(tf_idf == max(tf_idf)) %>% 
  ungroup() %>% 
  slice_max(tf_idf, n=50)

head(top50_terms)
```

